---
title: "Draft Preliminary Report on the Survey for the Revision of the SAA's Principles of Archaeological Ethics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
  number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 2)
```

```{r load-pkgs}
library(tidyverse)
library(magrittr)
library(likert)
library(knitr)
library(UpSetR)
library(patchwork)

# load our custom functions
source(here::here("analysis/functions.R"))
load_survey_data()
decode_demographics()
```

# Demographics of respondants {-}

Are you a current SAA member?	

```{r}
library(janitor)
survey_data %>% 
  select(survey_questions_vec[single_response[1]]) %>% 
  tabyl(`Are you a current SAA member?`) %>% 
  adorn_totals("row")
```

What is your primary work setting?	

```{r work-setting}
work_setting <- 
  plot_single_option_question(survey_questions_vec[single_response[11]], 
                                 base_size = 10,
                                  wrap = 100,
                                 title_wrap = 50,
                                 sort_bars = TRUE)

p4 <- 
  work_setting

library(janitor)
survey_data %>% 
  select(survey_questions_vec[single_response[11]]) %>% 
  mutate(simpler = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
   str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
   str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
   TRUE ~ "other"
  )) %>% 
  tabyl(simpler ) %>% 
  adorn_totals("row")
```


Is SAA membership correlated with certain work settings?

```{r}
library(janitor)
SAA_member_and_work_settings <- 
survey_data %>% 
  select(survey_questions_vec[single_response[11]],
         survey_questions_vec[single_response[1]]) %>% 
  mutate(simpler = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
   str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
   str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
   TRUE ~ "other"
  )) %>% 
  drop_na %>% 
  tabyl(`Are you a current SAA member?`, simpler)

chisq.test(SAA_member_and_work_settings)$residuals

```

Are you a member of the RPA?	

```{r}
#  member of RPA?
library(janitor)
survey_data %>% 
  select(survey_questions_vec[single_response[4]]) %>% 
  tabyl(`Are you a member of the RPA?`) %>% 
  adorn_totals("row")
```

Are RPA members more likely to be working in CRM?

```{r}
# RPA and CRM?
RPA_and_CRM_tbl <- 
survey_data %>% 
  select(survey_questions_vec[single_response[4]],
         survey_questions_vec[single_response[11]]) %>% 
   mutate(simpler = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
   str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
   str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
   TRUE ~ "other"
  )) %>% 
  select(`Are you a member of the RPA?`, simpler) %>% 
  #group_by(`Are you a member of the RPA?`, simpler) %>% 
  #tally %>% 
  drop_na() %>% 
  tabyl(`Are you a member of the RPA?`, simpler)

RPA_and_CRM_tbl

chisq.test(RPA_and_CRM_tbl)
```

How many years worked in archaeology?

```{r how-many-years, fig.height=3}
# how many years worked in archaeology?
how_many_years_plot <- 
survey_data %>% 
  select(`How many years have you worked in/with archaeology?`) %>% 
  ggplot(aes(`How many years have you worked in/with archaeology?`)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = mean(survey_data$`How many years have you worked in/with archaeology?`, na.rm = TRUE),
             colour = "red") +
  theme_bw()

library(ggstance)
how_many_years_plot <- 
survey_data %>% 
  select(`How many years have you worked in/with archaeology?`) %>% 
  ggplot(aes(`How many years have you worked in/with archaeology?`)) +
  geom_histogram(binwidth = 5) +
  geom_boxploth(aes(y = -20), width = 20) +
  theme_bw()
```

How many years have you worked in/with archaeology?	

Any relationship between work setting and years in archaeology?

```{r}
library(janitor)
years_and_work_settings <- 
survey_data %>% 
  select(survey_questions_vec[single_response[11]],
         `How many years have you worked in/with archaeology?`) %>% 
  mutate(simpler = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
   str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
   str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
   TRUE ~ "other"
  )) %>% 
  drop_na %>% 
  select(`How many years have you worked in/with archaeology?`, simpler)

ggplot(years_and_work_settings) +
  aes(simpler, 
      `How many years have you worked in/with archaeology?`) +
  geom_boxplot() +
  theme_bw()

# any significant difference? no
broom::tidy(aov(`How many years have you worked in/with archaeology?` ~ simpler ,
                data = years_and_work_settings))

```

Where do you do fieldwork/research? 

```{r}
# Where do you do fieldwork/research? 
where_do_you_do_fieldwork <- 
  plot_multiple_option_question(survey_questions_vec[multiple_response[4]],
                                wrap = 50,
                                truncn = 100,
                            title_wrap = 40)

p3 <- where_do_you_do_fieldwork$plot
```


```{r fig.height=6, fig.width=10}
library(patchwork)
wrap_plots(p3,  p4, ncol = 1)

ggsave(here::here("figures/demographics-work-fieldwork.png"),
                  height = 10,
                  width = 10)
```


```{r}
# Where do you do fieldwork/research? 
library(janitor)
survey_data %>% 
  select(survey_questions_vec[multiple_response[4]]) %>% 
  separate_rows(`Where do you do fieldwork/research? Please check all that apply.` , sep = ",") %>% 
  mutate(where_fieldwork = tolower(str_squish(`Where do you do fieldwork/research? Please check all that apply.`))) %>% 
  tabyl(where_fieldwork) %>% 
  adorn_totals("row") %>% 
  arrange(-n)
```

## Demo vars that were anonymised

Geographical area of origin

```{r}
library(janitor)
geoorigin_tally <- 
survey_data_demographics %>% 
  tabyl(geoorigin) %>% 
  arrange(-n) %>% 
  rename(variable = geoorigin)

ggplot(geoorigin_tally) +
  aes(variable, n) +
  geom_col() +
  theme_bw()
```

Current place of residence 

```{r}
residence_tally <- 
survey_data_demographics %>% 
  tabyl(residence) %>% 
  arrange(-n)  %>% 
  rename(variable = residence)

ggplot(residence_tally) +
  aes(variable, n) +
  geom_col() +
  theme_bw()
```

Age

```{r}
age_fct <- c("18-22", "23-29", "30-39", "40-49", "50-59", "60-69", "70 and over", "Prefer not to answer")

age_tally <- 
survey_data_demographics %>% 
  mutate(age = factor(age, levels = age_fct)) %>% 
  tabyl(age) %>% 
  arrange(age) %>% 
  rename(variable = age)

ggplot(age_tally) +
  aes(variable, n) +
  geom_col() +
  theme_bw()
```

Gender

```{r}
gender_tally <- 
survey_data_demographics %>% 
  tabyl(gender) %>% 
  arrange(-n) %>% 
  rename(variable = gender)

ggplot(gender_tally) +
  aes(variable, n) +
  geom_col() +
  theme_bw()
```

LGBTQIA+

```{r}
LGBTQIA_tally <- 
survey_data_demographics %>% 
  tabyl(lgbt_comm) %>% 
  arrange(-n) %>% 
  rename(variable = lgbt_comm)

ggplot(LGBTQIA_tally) +
  aes(variable, n) +
  geom_col() +
  theme_bw()
```

Ethnicity

```{r}
ethnicity_tally <- 
survey_data_demographics %>% 
  tabyl(ethnicity) %>% 
  arrange(-n) %>% 
  rename(variable = ethnicity)

ggplot(ethnicity_tally) +
  aes(variable, n) +
  geom_col() +
  theme_bw()
```

Combine some demo vars

```{r}
dv <- 
bind_rows("Geographic origin" = geoorigin_tally,
          "Country of residence" = residence_tally,
          "Age" = age_tally,
          "Gender" = gender_tally,
          "LGBTQIA+" = LGBTQIA_tally,
          "Ethnicity" = ethnicity_tally,
          .id = "var")

ggplot(dv) +
  aes(variable, n) +
  geom_col() +
  facet_wrap(~ var,
             scales = "free",
             ncol = 1) +
  theme_bw(base_size = 8) +
  scale_x_discrete(guide=guide_axis(n.dodge=2)) +
  xlab("")
```



Summary multi-panel plot with most demographic variables

```{r}

survey_data_demographics_member <- 
survey_data_demographics %>% 
  bind_cols(saa = survey_data$`Are you a current SAA member?`,
            rpa = survey_data$`Are you a member of the RPA?`)

plot_labels <- 
  c("Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin",
    "SAA member",
    "RPA member"
  )


horizontal_demo_plots <- 
    map2(survey_data_demographics_member, 
         plot_labels,
    ~tabyl(.x) %>% 
       ggplot( 
           aes(str_wrap(.x, 50),
               n)) +
      geom_col() +
      coord_flip() +
      xlab("") +
      scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) +
      theme_bw(base_size = 12) +
      ggtitle(.y)
    )
      
```


```{r demographics, fig.height=10}
library(cowplot)
p01 <- plot_grid(plotlist = horizontal_demo_plots, 
                 ncol = 2,
                 axis = "lr",
                 align = "hv")

plot_grid(p01,
          how_many_years_plot,
          ncol = 1,
          rel_heights = c(1, 0.3))


ggsave(here::here("figures/demographics-summary-panel.png"),
                  height = 15,
                  width = 15)
```

MCA to explore structure in the demographics of the respondents 

```{r}
# demo vars
demographic_vars <- 
  survey_data[, c(45:47, 49:50, 52:53, 55) ] %>% 
  bind_cols(survey_data_demographics_member) %>% 
    mutate(`Work setting` = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
   str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
   str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
   TRUE ~ "other"
  )) %>% 
  dplyr::select(-`What is your age?`,
                -`What is your gender identity? - Selected Choice`,
                -`Do you consider yourself a member of the LGBTQIA+ community?`,
                -`Please indicate your ethnicity: - Selected Choice`,
                -`What is your primary work setting? - Selected Choice`,
                -`How many years have you worked in/with archaeology?`,
                -`Current place of residence:`,
                -`Geographical area of origin:`
              ) %>% 
  mutate_all(as.factor) %>% 
  filter_all(all_vars(!str_detect(., "Prefer|other|Other"))) %>% 
  drop_na

names(demographic_vars) <-  c(plot_labels, "Work setting")

demographic_vars %<>% 
  select(-`Country of residence`,
         -`Geographic origin`)

cats = apply(demographic_vars, 2, function(x) nlevels(as.factor(x)))

require(FactoMineR)
require(ggplot2)

# apply MCA
mca1 = MCA(as.data.frame(demographic_vars), 
           graph = FALSE)

# list of results
mca1

# data frame with variable coordinates
mca1_vars_df = 
  data.frame(mca1$var$coord,
             Variable = rep(names(cats), cats))

# data frame with observation coordinates
mca1_obs_df = data.frame(mca1$ind$coord)

# tidy row names 
rownames(mca1_vars_df) <- str_replace(rownames(mca1_vars_df), "_", ": ")
rownames(mca1_vars_df) <- ifelse(str_detect(rownames(mca1_vars_df), ": "),
                                 rownames(mca1_vars_df),
                                 paste0(mca1_vars_df$Variable, ": ", rownames(mca1_vars_df)))
rownames(mca1_vars_df) <- str_remove(rownames(mca1_vars_df), "\\.$")
rownames(mca1_vars_df) <- 
  str_replace(rownames(mca1_vars_df), "African-American", "African American" )
                                 

# plot of variable categories
ggplot(data=mca1_vars_df, 
       aes(x = Dim.1, y = Dim.2, label = rownames(mca1_vars_df))) +
 geom_hline(yintercept = 0, colour = "gray70") +
 geom_vline(xintercept = 0, colour = "gray70") +
 geom_text(aes(colour=Variable)) +
 ggtitle("MCA plot of variables using R package FactoMineR")

# MCA plot of observations and categories
library(ggrepel)
ggplot(data = mca1_obs_df, 
       aes(x = Dim.1, y = Dim.2)) +
  geom_hline(yintercept = 0, colour = "gray70") +
  geom_vline(xintercept = 0, colour = "gray70") +
  geom_point(colour = "gray50", alpha = 0.7) +
  #geom_density2d(colour = "gray80") +
  geom_text_repel(data = mca1_vars_df, 
            aes(x = Dim.1, 
                y = Dim.2, 
                label = rownames(mca1_vars_df), 
                colour = Variable),
            size = 3,
            force = 10,
            bg.r = 0.1,
            bg.color = "white",
            max.overlaps = 20) +
  scale_colour_discrete(name = "Variable") +
  theme_bw(base_size = 12) +
  theme(legend.position = c(0.85, 0.7))

ggsave(here::here("figures/demographics-mca.jpg"),
       width = 6,
       height = 6,
       dpi = 600)

```




