---
title: "Draft Preliminary Report on the Survey for the Revision of the SAA's Principles of Archaeological Ethics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
  number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 2)
```

```{r load-pkgs}
library(tidyverse)
library(magrittr)
library(likert)
library(knitr)
library(UpSetR)
library(patchwork)

# load our custom functions
source(here::here("analysis/functions.R"))
load_survey_data()
decode_demographics()
```


# As a whole, the SAA Principles of Ethics satisfactorily addresses the ethical situations archaeologists find themselves in today

```{r}
situations <- 
survey_data %>% 
  select(`As a whole, the SAA Principles of Ethics satisfactorily addresses the ethical situations archaeologists find themselves in today.`,
         `Please elaborate here:_9`,
         `The SAA Principles of Ethics adequately addresses ethical concerns in the country in which I work, am a student, or teach.`,
        `Please elaborate here:_10`,
        `The SAA Principles of Ethics adequately addresses ethical concerns in sector in which I work.`,
        `Please elaborate here:_11`)
```


```{r}
# likert scale

library(grid)
  mylevels <- c('Strongly Disagree', 'Disagree', 'Neutral', 'Agree', 'Strongly Agree')
  
output <- 
  situations %>% 
  select(-contains("Please")) %>% 
  drop_na() %>% 
  mutate_all(~factor(., levels=mylevels)) %>% 
  as.data.frame()

names(output) <- 
  c("Ethical situations today", 
    "Concerns in country", 
    "Concerns in sector")

library(likert)
l_out <- likert(output)
  
plot(l_out, include.histogram = FALSE) +
 theme_minimal(base_size = 15)

ggsave(here::here("figures/concerns-situations-likert.jpg"),
       width = 10,
       height = 2,
       dpi = 600)
```

How many people responded to these questions about the situations and concerns? `r nrow(output)` people. Here's a table of how many responses in each category for each question:

```{r}
output_long <- 
  output %>% 
  pivot_longer( cols = everything(),
                names_to = "question", 
                values_to = "response") %>% 
  group_by(question, response) %>% 
  tally() %>% 
  pivot_wider(names_from = response, 
              values_from = n)  %>% 
  select(rev(mylevels)) %>% 
  janitor::adorn_totals("col")

write.csv(output_long,
          row.names = FALSE,
          here::here("data/derived-data/situations-and-concerns-likert-tally.csv"))
```

Can we get a summary of average disagreement for each demographic for each Q?

```{r}

survey_data_likert_demo_situation <- 
bind_cols(survey_data_demographics, 
          survey_data %>% 
  select(`As a whole, the SAA Principles of Ethics satisfactorily addresses the ethical situations archaeologists find themselves in today.`,
         `The SAA Principles of Ethics adequately addresses ethical concerns in the country in which I work, am a student, or teach.`,
        `The SAA Principles of Ethics adequately addresses ethical concerns in sector in which I work.`,
                 rpa = `Are you a member of the RPA?`)) %>% 
  relocate(rpa, .after = geoorigin) %>% 
  drop_na() %>% 
  filter_all(all_vars(!str_detect(., 
                                  "Prefer|other|Other")))  %>% 
  # high number == more disagree
  mutate_at(vars(`As a whole, the SAA Principles of Ethics satisfactorily addresses the ethical situations archaeologists find themselves in today.`,
         `The SAA Principles of Ethics adequately addresses ethical concerns in the country in which I work, am a student, or teach.`,
        `The SAA Principles of Ethics adequately addresses ethical concerns in sector in which I work.`), 
            ~fct_rev(factor(., levels=mylevels)) ) %>% 
  mutate_at(vars(`As a whole, the SAA Principles of Ethics satisfactorily addresses the ethical situations archaeologists find themselves in today.`,
         `The SAA Principles of Ethics adequately addresses ethical concerns in the country in which I work, am a student, or teach.`,
        `The SAA Principles of Ethics adequately addresses ethical concerns in sector in which I work.`), 
            ~as.numeric(.) ) 
  

# get col names
demo_cols <- c(names(survey_data_likert_demo_situation)[1:7])
names(demo_cols) <-  
  c("Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin",
    "RPA member"
  )
p_cols <- names(survey_data_likert_demo_situation)[8:ncol(survey_data_likert_demo_situation)]

mean_diss_out <- vector("list", length = length(p_cols))
for(i in 1:length(p_cols)){

demo_means <- 
map(demo_cols,
  ~bind_cols(
  survey_data_likert_demo_situation[.x],
  survey_data_likert_demo_situation[p_cols[i]]
   ) %>% 
  group_by(!!as.name(.x)) %>% 
  summarise(mean_disagreement = mean(!!as.name(p_cols[i])),
            n = n()) %>% 
  rename(var = .x)
 ) %>% 
  bind_rows(.id = "demo_var")

demo_values <- 
map(demo_cols,
  ~bind_cols(
  survey_data_likert_demo_situation[.x],
  survey_data_likert_demo_situation[p_cols[i]]
   ) %>% 
  rename(var = .x) 
 ) %>% 
  bind_rows(.id = "demo_var")

# https://bookdown.org/Rmadillo/likert/is-there-a-significant-difference.html
library(MASS)
library(ordinal)
library(bbmle)
library(coin)

# do a stat test for differences
xy <- 
demo_values %>% 
  mutate(li = factor(!!as.name(p_cols[i]))) %>% 
  mutate(var_fct = factor(var)) %>% 
  nest_by(demo_var) %>% 
  mutate(e_test = list(broom::tidy(kruskal.test(as.numeric(li) ~ factor(var), data = data)))) %>% 
  unnest(e_test)  %>% 
  unnest(data) %>% 
  right_join(demo_means) %>% 
  mutate(demo_var_test = paste0(demo_var, 
                                " (H(", parameter, ") = ", 
                                round(statistic, 2),
                                ", p = ", round(p.value, 3) )) 

shift_legend3 <- function(p) {
    pnls <- cowplot::plot_to_gtable(p) %>% gtable::gtable_filter("panel") %>%
      with(setNames(grobs, layout$name)) %>% purrr::keep(~identical(.x,zeroGrob()))

    if( length(pnls) == 0 ) stop( "No empty facets in the plot" )

    lemon::reposition_legend( p, "center", panel=names(pnls) )
}

mean_dis_plot <-         
ggplot(xy) +
  aes(var, 
      mean_disagreement) +
  geom_jitter(data = xy, 
                   aes(var, 
                       !!as.name(p_cols[i])),
                   alpha = 0.1
                   ) +
    geom_point(colour = "red",
               stroke = 2,
               aes(size = n,
                   shape = ifelse(p.value < 0.05, 16, 1))) +
  scale_shape_identity() +
  facet_wrap( ~ demo_var_test,
              scales = "free") +
  scale_y_continuous(limits = c(0,5), 
                     name = "Mean disagreement") +
  xlab("") +
  coord_flip() +
  theme_bw()

mean_diss_out[[i]] <- shift_legend3(mean_dis_plot)


ggsave(
  here::here(paste0("figures/concerns-situations-mean-dis-", i, ".png")),
  plot = mean_diss_out[[i]] ,
       width = 15,
       height = 10)
}
```

How many free text responses for each sitation and concern question?

Network plots for the high frequency words in the free text responses

```{r}
situations_elaboration <- 
situations %>% 
  dplyr::select(contains("Please")) %>% 
   rename_all(~names(output) ) %>% 
  mutate(across(everything(), stringi::stri_trans_tolower))

# count non-NA values for each principle 
p_count_situations_elaboration_responses <- 
  map_df(situations_elaboration, 
      ~length(.x[!is.na(.x)]))  %>% 
    pivot_longer(
      cols = everything(),
      names_to = 'question',
      values_to = 'responses')
  
library(knitr)
library(kableExtra)
library(formattable)

p_count_situations_elaboration_responses$responses <- 
  #color_bar("lightblue")(p_count_free_text_responses$responses)
color_tile("transparent", "red")(p_count_situations_elaboration_responses$responses)

kbl(p_count_situations_elaboration_responses, escape = F) %>% 
  column_spec(2, width = "1cm") 
```



```{r}

# Word cloud
library(quanteda)

nw <- vector("list", length = ncol(situations_elaboration))
for(i in 1:ncol(situations_elaboration)){
  
# i <- 3
x <- situations_elaboration %>% pull(i)

x_corp <- corpus(x)

set.seed(10)
dfmat1 <- dfm(x_corp,
              remove = c(stopwords("english"), 
                         "archaeological",
                         "archaeologists",
                         "archaeology",
                         "principle",
                         "principles",
                         "ethics",
                         "ethical",
                         "sector",
                         "country",
                         "countries",
                         "concerns",
                         "address",
                         "saa"), 
              remove_punct = TRUE,
              remove_numbers = TRUE) %>%
   dfm_trim(min_termfreq = 5) 

df_fcm <- fcm(dfmat1)
top_fcm <- 
  fcm_select(df_fcm, 
             pattern = names(quanteda::topfeatures(dfmat1, n = 20)))
nw[[i]] <- textplot_network(top_fcm, 
                       min_freq = 0.75, 
                       edge_alpha = 0.3, 
                       edge_size = 5)
}



library(cowplot)
plot_grid(plotlist = nw)

# nw[[1]]
for(i in 1:length(nw)){

ggsave(plot = nw[[i]],
       here::here(paste0("figures/concerns-situations-network-",
                         i, 
                         ".png")),
       width = 5,
       height = 5)

}

```


```{r}
## Average disagreement by work setting category
x1 <- 
survey_data %>% 
  # get demographic data
  mutate(work_setting = as.factor(case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
    str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
    str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
    TRUE ~ "other")
  ),
  years = `How many years have you worked in/with archaeology?`,
  member = `Are you a current SAA member?`,
  member = ifelse(str_detect("Yes", member), 
                  "yes", 'no'),
  rpa = `Are you a member of the RPA?`,
  age = `What is your age?`,
  gender = `What is your gender identity? - Selected Choice`,
  lgbqtia = `Do you consider yourself a member of the LGBTQIA+ community?`,
  ethnicity = `Please indicate your ethnicity: - Selected Choice`
  ) %>% 
  # get likert data for situations
  mutate_at(vars(29, 31, 33), ~factor(., levels=mylevels))  %>% 
  dplyr::select(
    29:34,
         work_setting
         ) %>% 
  drop_na() 

# cols 29, 31, 33
zz1 <- as.numeric(fct_rev(x1[[ names(survey_data)[29] ]] ))
zz2 <- as.numeric(fct_rev(x1[[ names(survey_data)[31] ]] ))
zz3 <- as.numeric(fct_rev(x1[[ names(survey_data)[33] ]] ))

x2 <- 
x1 %>% 
  dplyr::select(work_setting, 
                # years, 
                # member, 
                # rpa, 
                # age
                )  %>% 
  mutate_all(as.character) %>% 
  bind_cols(`addresses situations` = zz1,
            `concerns in country` = zz2,
            `concerns in sector` = zz3) %>% 
  pivot_longer(c(`addresses situations`, 
               `concerns in country`, 
               `concerns in sector`)) %>% 
  group_by(work_setting, name) %>% 
  summarise(`Average disagreement` = mean(value)) 


# average disagreement by with situation/country/sector by work setting
  ggplot(x2) +
  aes(work_setting,
      `Average disagreement`) +
  geom_col() +
    facet_wrap( ~ name)  +
  scale_y_continuous(limits = c(0, 5)) +
  theme_bw(base_size = 12) +
  xlab("")
  
ggsave(here::here("figures/concerns-situations-disagreemnt-panel.png"),
                  height = 5,
                  width = 10)
  

```









