---
title: "Draft Preliminary Report on the Survey for the Revision of the SAA's Principles of Archaeological Ethics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
  number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 2)
```

```{r load-pkgs}
library(tidyverse)
library(magrittr)
library(likert)
library(knitr)
library(UpSetR)
library(patchwork)

# load our custom functions
source(here::here("analysis/functions.R"))
load_survey_data()
decode_demographics()
```


Q: What do you see as being the primary ethical concerns in the field today?


```{r}
primary_concerns_demo <- 
survey_data %>% 
  mutate(work_setting = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
    str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
    str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
    TRUE ~ "other")
  ) %>% 
  select(`What do you see as being the primary ethical concerns in the field today?`,
         work_setting) %>% 
  bind_cols(survey_data_demographics)  %>% 
  drop_na() %>% 
  filter_all(all_vars(!str_detect(., 
                                  "Prefer|other|Other"))) 
```

primary concerns keyword network plot

```{r}
library(quanteda)
x_corp <- corpus(primary_concerns_demo, 
                 text_field = "What do you see as being the primary ethical concerns in the field today?")

set.seed(10)
dfmat1 <- dfm(x_corp,
              remove = c(stopwords("english"), 
                         "archaeological",
                         "archaeologists",
                         "archaeology",
                         "principle",
                         "ethics",
                         "ethical",
                         "saa"), 
              remove_punct = TRUE,
              remove_numbers = TRUE) %>%
   dfm_trim(min_termfreq = 5) 

# drop the empty docs
dfmat1  <- 
  dfm_subset(dfmat1, ntoken(dfmat1) > 0)

df_fcm <- fcm(dfmat1)
top_fcm <- 
  fcm_select(df_fcm, 
             pattern = names(quanteda::topfeatures(dfmat1, n = 20)))
nw <- textplot_network(top_fcm, 
                       min_freq = 0.75, 
                       edge_alpha = 0.3, 
                       edge_size = 5)

ggsave(plot = nw,
       here::here("figures/primary-ethical-concerns-network.png"),
       width = 5,
       height = 5)


```

Primary concerns topic model and diagnostics

```{r}
library(stm)
set.seed(12)

stm_out <- 
  stm(dfmat1, 
      seed = 12,
      init.type = "Spectral", 
      K = 0, # 49 topics
      prevalence = ~ work_setting + gender + age, 
      data = docvars(dfmat1),
      verbose = TRUE) 

library(tidytext)
td_beta <- tidy(stm_out)
td_gamma <- tidy(stm_out, 
                 matrix = "gamma",
                 document_names = rownames(dfmat1))

library(ggthemes)

top_terms <- td_beta %>%
  arrange(beta) %>%
  group_by(topic) %>%
  top_n(7, beta) %>%
  arrange(-beta) %>%
  select(topic, term) %>%
  summarise(terms = list(term)) %>%
  mutate(terms = map(terms, paste, collapse = ", ")) %>% 
  unnest()

gamma_terms <- td_gamma %>%
  group_by(topic) %>%
  summarise(gamma = mean(gamma)) %>%
  arrange(desc(gamma)) %>%
  left_join(top_terms, by = "topic") %>%
  mutate(topic = paste0("Topic ", topic),
         topic = reorder(topic, gamma))

tm <- 
gamma_terms %>%
  top_n(20, gamma) %>%
  ggplot(aes(topic, 
             gamma, 
             label = str_wrap(terms, 60), 
             fill = topic)) +
  geom_col(show.legend = FALSE) +
  geom_text(hjust = 0, 
            vjust = 0.5,
            lineheight = .75,
            nudge_y = 0.0005, 
            size = 2.5) +
  scale_fill_viridis_d(guide = "legend") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, 0.09),
                     labels = scales::percent_format()) +
  labs(x = NULL, y = expression(gamma),
       title = "Top 20 topics by prevalence in the question\n'What do you see as being the primary ethical concerns in the field today?'",
       subtitle = "With the top words that contribute to each topic") +
  theme_bw()

# https://francescocaberlin.blog/2019/07/01/messing-around-with-stm-part-iiib-model-analysis/
ExSem <-
  as.data.frame(cbind(
    k = c(1:stm_out$settings$dim$K),
    exclusivity(stm_out),
    semanticCoherence(model = stm_out, 
                      dfmat1)
  ))

colnames(ExSem)<-c("k", "Exclusivity", "SemanticCoherence")

ExSem_gamma <- 
gamma_terms %>% 
  mutate(k = parse_number(as.character(topic))) %>% 
  left_join(ExSem)

library(ggrepel)
plotexcoer <-
  ggplot(ExSem_gamma, 
         aes(SemanticCoherence, 
             Exclusivity,
             colour = gamma,
             size = gamma)) +
  geom_point(alpha = 0.7) +
  geom_text_repel(aes(label = k)) +
  labs(x = "Semantic coherence",
       y = "Exclusivity",
       title = "Comparing topic exclusivity and topic semantic coherence") +
  scale_colour_viridis_c(guide = "legend") +
  theme_bw()

library(cowplot)
plot_grid(
  tm, 
  plotexcoer,
  align = "h"
  )

ggsave(here::here('figures/primary-ethical-concerns-topic-model.png'),
       height = 7,
       width = 15)
```

Primary concerns topic model: topics and covariates

```{r}

K_est <- stm_out$settings$dim$K

topics_of_interest <- 
  ExSem_gamma %>% 
  arrange(desc(gamma)) %>% 
  slice(1:3, 11) %>% 
  pull(k)

topics_of_interest_terms <- 
    ExSem_gamma %>% 
  arrange(desc(gamma)) %>% 
  slice(1:3, 11) %>% 
  pull(terms)

# get vec of covars
covars <- names(primary_concerns_demo)[2:ncol(primary_concerns_demo)]

covars_pretty_names <- 
  c("Work setting",
    "Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin"
  )

for(j in seq_along(topics_of_interest_terms)){

covars_output <- vector("list", length = length((covars)))

for(i in seq_along(covars_output)){
prep <- estimateEffect(as.formula(paste("1:K_est ~", covars[i])),
                       stm_out,
                       docvars(dfmat1) %>% 
                         mutate_all(as.factor),
                       uncertainty = "Global")

plot_output <- 
plot.estimateEffect(
  prep,
  model = stm_out,
  covariate = covars[i],
  topics = topics_of_interest[j],
  method = "pointestimate",
  xlim = c(-.25, 0.25),
  omit.plot = TRUE
)

Result_tbl <-
tibble(means = unlist(plot_output$means),
       topics = unlist(plot_output$topics),
       labels = unlist(plot_output$labels),
       ci_lower  = unname(unlist(data.frame(plot_output$ci)[1,])),
       ci_upper = unname(unlist(data.frame(plot_output$ci)[2,]))) %>% 
  arrange(desc(means)) %>% 
  mutate(order = 1:n())


highlight_tbl <-
  Result_tbl %>%
  filter(ci_lower > 0 | ci_upper < 0)

normal_tbl <-
  Result_tbl %>%
  filter(ci_lower < 0) %>%
  filter(ci_upper > 0)

highlight_size = 2.2
normal_size = 2.2
normal_colour = "grey30"

covars_output[[i]] <-
ggplot() +
  theme_bw() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  geom_vline(xintercept = 0,
             colour = "red") +
  geom_linerange(data = normal_tbl,
               aes(y = order,
                   xmin =ci_lower,
                   xmax = ci_upper),
               colour = normal_colour) +
  geom_linerange(data = highlight_tbl,
                 aes(y = order,
                     xmin = ci_lower,
                     xmax = ci_upper),
                 colour = "red") +
  geom_linerange(data = highlight_tbl,
                 aes(y = order,
                     xmin =ci_lower,
                     xmax = ci_upper),
                 colour = "red") +
  geom_point(data = normal_tbl,
             aes(means,
                 order),
             colour = normal_colour) +
  geom_point(data = highlight_tbl,
             aes(means,
                 order),
             pch = 17,
             colour = "red",
             size = 2) +
  geom_text(data = normal_tbl,
            aes(means,
                order + 0.5,
                label = labels),
            size = normal_size) +
  geom_text(data = highlight_tbl,
            aes(means,
                order + 0.5,
                label = labels),
            size = highlight_size,
            colour = "red") +
  xlab("Estimate") +
  ylab(covars_pretty_names[i]) 

}

library(cowplot)


plot_grid( 
          plotlist = covars_output
) +
  annotate("text", 
           x = 0.4,
           y = 0.2,
           size = 6,
           hjust = 0,
           label = paste0("Topic ", 
                          topics_of_interest[j], 
                          ": ",
                          str_wrap(topics_of_interest_terms[j], 40)))


ggsave(here::here(paste0('figures/primary-ethical-concerns-topic-model-', 
                         topics_of_interest[j],
                         '-covars.png')),
       height = 8,
       width = 8)
       
}


```

Q: What should be the purpose of the SAA's ethical document?  

```{r}
purpose_options <- c(
"Establish best practices for the profession",
"Establish and enforce best practices for the profession",
"Establish expectations for the behavior of professionals",
"Demonstrate integrity of profession",
"For use in education and training")

purpose_of_ethics_doc <- 
survey_data %>% 
  mutate(work_setting = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
    str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
    str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
    TRUE ~ "other")
  ) %>% 
  select(starts_with("What should be the purpose of the SAA"),
         work_setting) %>% 
  bind_cols(survey_data_demographics)  %>% 
  drop_na() %>% 
  filter_all(all_vars(!str_detect(., 
                                  "Prefer|other|Other"))) %>% 
  drop_na() 
```

Purpose of doc ranking

```{r}
purpose_of_ethics_doc_wide_fn <- 
  function(x) {
  x %>% 
  set_names(str_c("rank ", 1:length(purpose_options))) %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = -id,
               names_to = "variable",
               values_to = "value") %>% 
  mutate(value = case_when(
    value == 1 ~ purpose_options[1],
    value == 2 ~ purpose_options[2],
    value == 3 ~ purpose_options[3],
    value == 4 ~ purpose_options[4],
    value == 5 ~ purpose_options[5],
    TRUE    ~ "other"
  )) 
  }

purpose_of_ethics_doc_wide <-  
  purpose_of_ethics_doc_wide_fn(purpose_of_ethics_doc[ , 1:5])

purpose_of_ethics_doc_wide$value <- 
  factor(purpose_of_ethics_doc_wide$value, levels = purpose_options)
```

Purpose of doc plots

```{r}
purpose_of_doc_plot <- 
purpose_of_ethics_doc_wide %>% 
  mutate(rank = parse_number(variable)) %>% 
  group_by(value) %>% 
  summarise(sum_ranks = sum(rank)) %>% 
  arrange(desc(sum_ranks)) %>% 
  ggplot() +
  aes(reorder(str_wrap(value, 20), 
              -sum_ranks),
      sum_ranks,
      fill = value) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  xlab("") +
  ylab("Sum of ranks (shorter bar indicates higher rank)") +
  scale_fill_viridis_d()  +
  guides(fill=FALSE)

purpose_of_doc_stacked_plot <- 
purpose_of_ethics_doc_wide %>% 
  mutate(rank = parse_number(variable)) %>% 
  ggplot() +
  aes(rank,
      id, 
      fill = value) +
   geom_bar(position = "fill", 
            stat = "identity") +
  theme_bw() +
  xlab("Rank") +
  ylab("Proportion") +
  scale_fill_viridis_d() +
  guides(fill=FALSE)

library(cowplot)
plot_grid(purpose_of_doc_plot, 
          purpose_of_doc_stacked_plot)

ggsave(here::here(paste0('figures/purpose-of-doc.png')),
       height = 6,
       width = 10)
       
```

Purpose of documents rankings by various demographic variables

```{r}
purpose_of_ethics_doc_wide_demo <- 
purpose_of_ethics_doc %>% 
  mutate(id = 1:n()) %>% 
  left_join(purpose_of_ethics_doc_wide) %>% 
    mutate(rank = parse_number(variable)) 

names(purpose_of_ethics_doc_wide_demo)[6:12] <- 
  c("Work setting",
    "Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin"
  )

# for each demo var
purpose_of_ethics_doc_wide_demo_wide <- 
purpose_of_ethics_doc_wide_demo %>% 
  pivot_longer(cols = `Work setting`:`Geographic origin`,
    names_to = "var1",
    values_to = "val1")

purpose_of_ethics_doc_wide_demo_wide_means <- 
purpose_of_ethics_doc_wide_demo_wide %>% 
  group_by(value, var1, val1) %>% 
  summarise(mean_rank = mean(rank),
            n = n()) 

purpose_of_ethics_doc_wide_demo_wide_means_plots <- 
purpose_of_ethics_doc_wide_demo_wide_means %>% 
  ungroup %>% 
  nest_by(var1) %>% 
  mutate(plot_out = list(ggplot(data = data) +
  aes(value, 
      mean_rank,
      group = val1,
      colour = val1) +
  geom_point(aes(size = n)) +
  geom_line() +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_color_viridis_d(name = var1) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_y_reverse() ))

library(cowplot)
plot_grid(plotlist = purpose_of_ethics_doc_wide_demo_wide_means_plots$plot_out,
          align = "hv")

ggsave(here::here(paste0('figures/purpose-of-doc-demographics.png')),
       height = 14,
       width = 19)

```

Q: What type of ethical document(s) do you feel would best serve the SAA membership? 
```{r}
### Responses to the question "What type of ethical document(s) do you feel would best serve the SAA membership?" 

# What type of ethical document(s) do you feel would best serve the SAA membership? 
what_type_of_doc <- 
  plot_multiple_option_question(survey_questions_vec[multiple_response[3]])
```

```{r what-type-document-plot, fig.cap=what_type_of_doc$upset_plot_caption, fig.height=3}

png(here::here("figures/what-type-doc-upset-plot.png"),
    width = 10,
    height = 5,
    units = "in",
    res = 300)
what_type_of_doc$upset_plot
dev.off()
```

Type of ethical document(s) by demographic group "What type of ethical document"

```{r}
type_of_ethical_doc <- 
survey_data %>% 
  mutate(work_setting = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
    str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
    str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
    TRUE ~ "other")
  ) %>% 
  select(starts_with("What type of ethical document"),
         work_setting) %>% 
  bind_cols(survey_data_demographics)  %>% 
  drop_na() %>% 
  filter_all(all_vars(!str_detect(., 
                                  "Prefer|other|Other"))) %>% 
  drop_na() 

names(type_of_ethical_doc) <- 
  c("Type of document",
    "Work setting",
    "Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin"
  )

type_of_ethical_doc_demo_plots <- 
type_of_ethical_doc %>% 
  pivot_longer(`Work setting`:`Geographic origin`) %>% 
  separate_rows(`Type of document`, sep = ",") %>% 
  mutate(`Type of document` = factor(`Type of document`, 
                                     levels = what_type_of_doc$tally$x)) %>% 
  group_by(name, value) %>% 
  mutate(n_in_demo_group = n()) %>% 
  group_by(name, value, `Type of document`) %>% 
  summarise(n_doc_type = n(),
            n = unique(n_in_demo_group)) %>% 
  mutate(freq_doc_type = n_doc_type / sum(n_doc_type))  %>% 
  ungroup() %>% 
  nest_by(name) %>% 
  mutate(type_plot = list(ggplot(data) +
                           aes(`Type of document`,
                           freq_doc_type,
                           colour = value,
                           group = value) +
                           geom_line() +
                           geom_point(aes(size = n)) +
                           xlab("") +
                           ylab("") +
                           theme_bw() +
              scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
              scale_color_viridis_d(name = name) +
              theme(axis.text.x = element_text(size = 6)) 
              )
         )

library(cowplot)
plot_grid(plotlist = type_of_ethical_doc_demo_plots$type_plot,
          align = "hv")

ggsave(here::here(paste0('figures/type-of-doc-demographics.png')),
       height = 14,
       width = 19)

type_age_corr <- 
type_of_ethical_doc %>% 
  pivot_longer(`Work setting`:`Geographic origin`) %>% 
  separate_rows(`Type of document`, sep = ",") %>% 
  mutate(`Type of document` = factor(`Type of document`, 
                                     levels = what_type_of_doc$tally$x)) %>% 
  filter(name == "Age") %>% 
  mutate(value = factor(value, levels = age_fct)) 

# years in archaeology by doc type
what_type_of_doc_by_year_worked <- 
  survey_data %>% 
  select(survey_questions_vec[multiple_response[3]],
         `How many years have you worked in/with archaeology?`) %>% 
  rename(what_type = survey_questions_vec[multiple_response[3]],
         years_worked = `How many years have you worked in/with archaeology?`) %>% 
  drop_na() %>% 
  separate_rows(what_type, 
                sep = ",(?=[[:upper:]])") %>% 
  group_by(what_type, 
           years_worked) %>% 
  tally() %>% 
  left_join(what_type_of_doc$tally,
            by = c("what_type" = "x"))

ggplot(what_type_of_doc_by_year_worked,
       aes(years_worked,
           n.x,
           colour = what_type)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(size = log10(n.y)), 
              se = FALSE) +
  scale_color_viridis_d(name = "Document type", 
                        alpha = 0.8) +
  theme_minimal(base_size = 14) +
  theme(
        legend.position = c(.8, 
                            .8),
        legend.box="vertical"
        ) +
  guides(colour=guide_legend(nrow=6, 
                             byrow=TRUE,
                             override.aes = list(size=5)),
         size = FALSE) +
  coord_cartesian(ylim = c(0, 15)) +
  xlab("Years worked in archaeology") +
  ylab("Count of responses")

ggsave(here::here(paste0('figures/type-of-doc-years-worked.png')),
       height = 7,
       width  = 12)
```


Q: Are there additional ethical issues that should be addressed by the Principles? 

```{r}
plot_single_option_question(survey_questions_vec[single_response[3]], 
                                 base_size = 12,
                                  wrap = 60,
                                 title_wrap = 60,
                            sort_bars = TRUE)

ggsave(here::here(paste0('figures/additional-ethical-issues-plot.png')),
       height = 3,
       width  = 7)


# percentages
library(janitor)
additional_ethical_issues_tally <- 
survey_data %>% 
  select(q = survey_questions_vec[single_response[3]]) %>% 
  group_by(q) %>% 
  tabyl(q)
```

```{r}
additional_ethical_issues <- 
survey_data %>% 
  mutate(work_setting = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
    str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
    str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
    TRUE ~ "other")
  ) %>% 
  select(starts_with("Are there additional ethical issues"),
         work_setting) %>% 
  bind_cols(survey_data_demographics)  %>% 
  drop_na() %>% 
  filter_all(all_vars(!str_detect(., 
                                  "Prefer|other|Other"))) %>% 
  drop_na() 

names(additional_ethical_issues) <- 
  c("Additional ethical issues",
    "Work setting",
    "Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin"
  )

# by dmographic


library(janitor)
additional_ethical_issues_tally %>% 
  ungroup() %>% 
  group_by(`Additional ethical issues`) %>% 
  tally %>% 
  mutate(perc = n / sum(n))


ggplot(additional_ethical_issues_tally) +
  aes(value, 
      n,
      fill = `Additional ethical issues`) +
  geom_col() +
  facet_wrap( ~ name, scales = "free") +
   scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    scale_fill_viridis_d(name = "") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 6),
                legend.position = c(.7, 
                                    .2))  +
  xlab("") +
  ylab("")

ggsave(here::here(paste0('figures/additional-ethical-issues-demo-plot.png')),
       height = 8,
       width  = 14)
```


Additional ethical issues - free text 

```{r}



# network plot ------------------------
additional <- 
survey_data %>% 
  select(survey_questions_vec[36]) %>% 
  drop_na()  %>% 
  pull() %>% 
  tolower() 

library(quanteda)
additional_cor <- corpus(additional)

set.seed(10)
dfmat2 <- dfm(additional,
              remove = c(stopwords("english"), 
                         "archaeological",
                         "archaeologists",
                         "archaeology",
                         "see",
                         "previous",
                         "comments",
                         "need",
                         "think",
                         "also",
                         "issues",
                         "issue",
                         "needs",
                         "can",
                         "saa",
                         "address",
                         "many",
                         "ethics",
                         "ethical",
                         "principle",
                         "principles",
                         "just",
                         "specific",
                         "really",
                         "much",
                         "better",
                         "another",
                         "mentioned",
                         "certain",
                         "like",
                         "probably",
                         "may",
                         "moment",
                         "additional"), 
              remove_punct = TRUE,
              remove_numbers = TRUE) %>%
   dfm_trim(min_termfreq = 3)

# basic wordcloud
wc2 <- ~{
  textplot_wordcloud(dfmat2, 
                   color = rev(RColorBrewer::brewer.pal(10, "RdBu")))
}

# word network
df_fcm2 <- fcm(dfmat2)
top_fcm2 <- fcm_select(df_fcm2, pattern = names(quanteda::topfeatures(dfmat2, n = 20)))
nw2 <- textplot_network(top_fcm2, 
                       min_freq = 0.75, 
                       edge_alpha = 0.3, 
                       edge_size = 5)


# topic model plot
library(stm)

stm_out2 <- 
  stm(dfmat2, 
      init.type="Spectral", 
      K = 0, # 45 topics
      verbose = FALSE) 

library(tidytext)
td_beta2 <- tidy(stm_out2)
td_gamma2 <- tidy(stm_out2, 
                 matrix = "gamma",
                 document_names = rownames(dfmat2))

library(ggthemes)

top_terms2 <- td_beta2 %>%
  arrange(beta) %>%
  group_by(topic) %>%
  top_n(7, beta) %>%
  arrange(-beta) %>%
  select(topic, term) %>%
  summarise(terms = list(term)) %>%
  mutate(terms = map(terms, paste, collapse = ", ")) %>% 
  unnest()

gamma_terms2 <- td_gamma2 %>%
  group_by(topic) %>%
  summarise(gamma = mean(gamma)) %>%
  arrange(desc(gamma)) %>%
  left_join(top_terms2, by = "topic") %>%
  mutate(topic = paste0("Topic ", topic),
         topic = reorder(topic, gamma))

tm2 <- 
gamma_terms2 %>%
  top_n(20, gamma) %>%
  ggplot(aes(topic, 
             gamma, 
             label = str_wrap(terms, 50), 
             fill = topic)) +
  scale_fill_viridis_d() +
  geom_col(show.legend = FALSE) +
  geom_text(hjust = 0, 
            vjust = 0.5,
            lineheight = .85,
            nudge_y = 0.0005, 
            size = 3) +
  coord_flip() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, 0.08),
                     labels = scales::percent_format()) +
  labs(x = NULL, y = expression(gamma),
       title = "Top 20 topics by prevalence in the question\n'Are there additional ethical issues that should be addressed by the Principles?'",
       subtitle = "With the top words that contribute to each topic") +
  theme_bw()


# https://francescocaberlin.blog/2019/07/01/messing-around-with-stm-part-iiib-model-analysis/
ExSem_add <-
  as.data.frame(cbind(
    k = c(1:stm_out2$settings$dim$K),
    exclusivity(stm_out2),
    semanticCoherence(model = stm_out2, 
                      dfmat2)
  ))

colnames(ExSem_add)<-c("k", "Exclusivity", "SemanticCoherence")

ExSem_add_gamma <- 
gamma_terms2 %>% 
  mutate(k = parse_number(as.character(topic))) %>% 
  left_join(ExSem_add)

library(ggrepel)
plotexcoer <-
  ggplot(ExSem_add_gamma, 
         aes(SemanticCoherence, 
             Exclusivity,
             colour = gamma,
             size = gamma)) +
  geom_point(alpha = 0.7) +
  geom_text_repel(aes(label = k),
                  force = 10) +
  labs(x = "Semantic coherence",
       y = "Exclusivity",
       title = "Comparing topic exclusivity and topic semantic coherence") +
  scale_colour_viridis_c(guide = "legend") +
  theme_bw() +
  theme(
        legend.position = c(.9, 
                            .2)
        )


library(cowplot)
library(gridGraphics)

dev.off()
# need to make the plot panel big to avoid some 'pin' error
plot_grid(plot_grid(wc2, nw2, 
                    ncol = 2,
                    labels=c("A","B")),
          plot_grid(tm2, plotexcoer,
                    labels=c("C","D"),
                    align = "hv"),
          nrow = 2,
          rel_heights = c(1, 2)) + 
  panel_border(remove = TRUE)

ggsave(here::here('figures/additional-ethical-issues-text-panel.png'),
       height = 12,
       width = 12)

dev.off()

```

























