---
title: "Draft Preliminary Report on the Survey for the Revision of the SAA's Principles of Archaeological Ethics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
  number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 2)
```

```{r load-pkgs}
library(tidyverse)
library(magrittr)
library(likert)
library(knitr)
library(UpSetR)
library(patchwork)

# load our custom functions
source(here::here("analysis/functions.R"))
load_survey_data()
decode_demographics()
```


Q: What do you see as being the primary ethical concerns in the field today?


```{r}
primary_concerns_demo <- 
survey_data %>% 
  mutate(work_setting = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
    str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
    str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
    TRUE ~ "other")
  ) %>% 
  select(`What do you see as being the primary ethical concerns in the field today?`,
         work_setting) %>% 
  bind_cols(survey_data_demographics)  %>% 
  drop_na() %>% 
  filter_all(all_vars(!str_detect(., 
                                  "Prefer|other|Other"))) 
```

primary concerns keyword network plot

```{r}
library(quanteda)
x_corp <- corpus(primary_concerns_demo, 
                 text_field = "What do you see as being the primary ethical concerns in the field today?")

set.seed(10)
dfmat1 <- dfm(x_corp,
              remove = c(stopwords("english"), 
                         "archaeological",
                         "archaeologists",
                         "archaeology",
                         "principle",
                         "ethics",
                         "ethical",
                         "saa"), 
              remove_punct = TRUE,
              remove_numbers = TRUE) %>%
   dfm_trim(min_termfreq = 5) 

# drop the empty docs
dfmat1  <- 
  dfm_subset(dfmat1, ntoken(dfmat1) > 0)

df_fcm <- fcm(dfmat1)
top_fcm <- 
  fcm_select(df_fcm, 
             pattern = names(quanteda::topfeatures(dfmat1, n = 20)))
nw <- textplot_network(top_fcm, 
                       min_freq = 0.75, 
                       edge_alpha = 0.3, 
                       edge_size = 5)

ggsave(plot = nw,
       here::here("figures/primary-ethical-concerns-network.png"),
       width = 5,
       height = 5)


```

Primary concerns topic model and diagnostics

```{r}
library(stm)
set.seed(12)

stm_out <- 
  stm(dfmat1, 
      seed = 12,
      init.type = "Spectral", 
      K = 0, # 49 topics
      prevalence = ~ work_setting + gender + age, 
      data = docvars(dfmat1),
      verbose = TRUE) 

library(tidytext)
td_beta <- tidy(stm_out)
td_gamma <- tidy(stm_out, 
                 matrix = "gamma",
                 document_names = rownames(dfmat1))

library(ggthemes)

top_terms <- td_beta %>%
  arrange(beta) %>%
  group_by(topic) %>%
  top_n(7, beta) %>%
  arrange(-beta) %>%
  select(topic, term) %>%
  summarise(terms = list(term)) %>%
  mutate(terms = map(terms, paste, collapse = ", ")) %>% 
  unnest()

gamma_terms <- td_gamma %>%
  group_by(topic) %>%
  summarise(gamma = mean(gamma)) %>%
  arrange(desc(gamma)) %>%
  left_join(top_terms, by = "topic") %>%
  mutate(topic = paste0("Topic ", topic),
         topic = reorder(topic, gamma))

tm <- 
gamma_terms %>%
  top_n(20, gamma) %>%
  ggplot(aes(topic, 
             gamma, 
             label = str_wrap(terms, 60), 
             fill = topic)) +
  geom_col(show.legend = FALSE) +
  geom_text(hjust = 0, 
            vjust = 0.5,
            lineheight = .75,
            nudge_y = 0.0005, 
            size = 2.5) +
  scale_fill_viridis_d(guide = "legend") +
  coord_flip() +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, 0.09),
                     labels = scales::percent_format()) +
  labs(x = NULL, y = expression(gamma),
       title = "Top 20 topics by prevalence in the question\n'What do you see as being the primary ethical concerns in the field today?'",
       subtitle = "With the top words that contribute to each topic") +
  theme_bw()

# https://francescocaberlin.blog/2019/07/01/messing-around-with-stm-part-iiib-model-analysis/
ExSem <-
  as.data.frame(cbind(
    k = c(1:stm_out$settings$dim$K),
    exclusivity(stm_out),
    semanticCoherence(model = stm_out, 
                      dfmat1)
  ))

colnames(ExSem)<-c("k", "Exclusivity", "SemanticCoherence")

ExSem_gamma <- 
gamma_terms %>% 
  mutate(k = parse_number(as.character(topic))) %>% 
  left_join(ExSem)

library(ggrepel)
plotexcoer <-
  ggplot(ExSem_gamma, 
         aes(SemanticCoherence, 
             Exclusivity,
             colour = gamma,
             size = gamma)) +
  geom_point(alpha = 0.7) +
  geom_text_repel(aes(label = k)) +
  labs(x = "Semantic coherence",
       y = "Exclusivity",
       title = "Comparing topic exclusivity and topic semantic coherence") +
  scale_colour_viridis_c(guide = "legend") +
  theme_bw()

library(cowplot)
plot_grid(
  tm, 
  plotexcoer,
  align = "h"
  )

ggsave(here::here('figures/primary-ethical-concerns-topic-model.png'),
       height = 7,
       width = 15)
```

Primary concerns topic model: topics and covariates

```{r}

K_est <- stm_out$settings$dim$K

topics_of_interest <- 
  ExSem_gamma %>% 
  arrange(desc(gamma)) %>% 
  slice(1:3, 11) %>% 
  pull(k)

topics_of_interest_terms <- 
    ExSem_gamma %>% 
  arrange(desc(gamma)) %>% 
  slice(1:3, 11) %>% 
  pull(terms)

# get vec of covars
covars <- names(primary_concerns_demo)[2:ncol(primary_concerns_demo)]

covars_pretty_names <- 
  c("Work setting",
    "Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin"
  )

for(j in seq_along(topics_of_interest_terms)){

covars_output <- vector("list", length = length((covars)))

for(i in seq_along(covars_output)){
prep <- estimateEffect(as.formula(paste("1:K_est ~", covars[i])),
                       stm_out,
                       docvars(dfmat1) %>% 
                         mutate_all(as.factor),
                       uncertainty = "Global")

plot_output <- 
plot.estimateEffect(
  prep,
  model = stm_out,
  covariate = covars[i],
  topics = topics_of_interest[j],
  method = "pointestimate",
  xlim = c(-.25, 0.25),
  omit.plot = TRUE
)

Result_tbl <-
tibble(means = unlist(plot_output$means),
       topics = unlist(plot_output$topics),
       labels = unlist(plot_output$labels),
       ci_lower  = unname(unlist(data.frame(plot_output$ci)[1,])),
       ci_upper = unname(unlist(data.frame(plot_output$ci)[2,]))) %>% 
  arrange(desc(means)) %>% 
  mutate(order = 1:n())


highlight_tbl <-
  Result_tbl %>%
  filter(ci_lower > 0 | ci_upper < 0)

normal_tbl <-
  Result_tbl %>%
  filter(ci_lower < 0) %>%
  filter(ci_upper > 0)

highlight_size = 2.2
normal_size = 2.2
normal_colour = "grey30"

covars_output[[i]] <-
ggplot() +
  theme_bw() +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  geom_vline(xintercept = 0,
             colour = "red") +
  geom_linerange(data = normal_tbl,
               aes(y = order,
                   xmin =ci_lower,
                   xmax = ci_upper),
               colour = normal_colour) +
  geom_linerange(data = highlight_tbl,
                 aes(y = order,
                     xmin = ci_lower,
                     xmax = ci_upper),
                 colour = "red") +
  geom_linerange(data = highlight_tbl,
                 aes(y = order,
                     xmin =ci_lower,
                     xmax = ci_upper),
                 colour = "red") +
  geom_point(data = normal_tbl,
             aes(means,
                 order),
             colour = normal_colour) +
  geom_point(data = highlight_tbl,
             aes(means,
                 order),
             pch = 17,
             colour = "red",
             size = 2) +
  geom_text(data = normal_tbl,
            aes(means,
                order + 0.5,
                label = labels),
            size = normal_size) +
  geom_text(data = highlight_tbl,
            aes(means,
                order + 0.5,
                label = labels),
            size = highlight_size,
            colour = "red") +
  xlab("Estimate") +
  ylab(covars_pretty_names[i]) 

}

library(cowplot)


plot_grid( 
          plotlist = covars_output
) +
  annotate("text", 
           x = 0.4,
           y = 0.2,
           size = 6,
           hjust = 0,
           label = paste0("Topic ", 
                          topics_of_interest[j], 
                          ": ",
                          str_wrap(topics_of_interest_terms[j], 40)))


ggsave(here::here(paste0('figures/primary-ethical-concerns-topic-model-', 
                         topics_of_interest[j],
                         '-covars.png')),
       height = 8,
       width = 8)
       
}


```

Q: What should be the purpose of the SAA's ethical document?  

```{r}
purpose_options <- c(
"Establish best practices for the profession",
"Establish and enforce best practices for the profession",
"Establish expectations for the behavior of professionals",
"Demonstrate integrity of profession",
"For use in education and training")

purpose_of_ethics_doc <- 
survey_data %>% 
  mutate(work_setting = case_when(
    str_detect(`What is your primary work setting? - Selected Choice`, "Academic") ~ "Academic",
    str_detect(`What is your primary work setting? - Selected Choice`, "CRM") ~ "CRM", 
    str_detect(`What is your primary work setting? - Selected Choice`, "Gov") ~ "Government",
    TRUE ~ "other")
  ) %>% 
  select(starts_with("What should be the purpose of the SAA"),
         work_setting) %>% 
  bind_cols(survey_data_demographics)  %>% 
  drop_na() %>% 
  filter_all(all_vars(!str_detect(., 
                                  "Prefer|other|Other"))) %>% 
  drop_na() 
```

Purpose of doc ranking

```{r}
purpose_of_ethics_doc_wide_fn <- 
  function(x) {
  x %>% 
  set_names(str_c("rank ", 1:length(purpose_options))) %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = -id,
               names_to = "variable",
               values_to = "value") %>% 
  mutate(value = case_when(
    value == 1 ~ purpose_options[1],
    value == 2 ~ purpose_options[2],
    value == 3 ~ purpose_options[3],
    value == 4 ~ purpose_options[4],
    value == 5 ~ purpose_options[5],
    TRUE    ~ "other"
  )) 
  }

purpose_of_ethics_doc_wide <-  
  purpose_of_ethics_doc_wide_fn(purpose_of_ethics_doc[ , 1:5])

purpose_of_ethics_doc_wide$value <- 
  factor(purpose_of_ethics_doc_wide$value, levels = purpose_options)
```

Purpose of doc plots

```{r}
purpose_of_doc_plot <- 
purpose_of_ethics_doc_wide %>% 
  mutate(rank = parse_number(variable)) %>% 
  group_by(value) %>% 
  summarise(sum_ranks = sum(rank)) %>% 
  arrange(desc(sum_ranks)) %>% 
  ggplot() +
  aes(reorder(str_wrap(value, 20), 
              -sum_ranks),
      sum_ranks,
      fill = value) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  xlab("") +
  ylab("Sum of ranks (shorter bar indicates higher rank)") +
  scale_fill_viridis_d()  +
  guides(fill=FALSE)

purpose_of_doc_stacked_plot <- 
purpose_of_ethics_doc_wide %>% 
  mutate(rank = parse_number(variable)) %>% 
  ggplot() +
  aes(rank,
      id, 
      fill = value) +
   geom_bar(position = "fill", 
            stat = "identity") +
  theme_bw() +
  xlab("Rank") +
  ylab("Proportion") +
  scale_fill_viridis_d() +
  guides(fill=FALSE)

library(cowplot)
plot_grid(purpose_of_doc_plot, 
          purpose_of_doc_stacked_plot)


ggsave(here::here(paste0('figures/purpose-of-doc.png')),
       height = 6,
       width = 10)
       
```

Purpose of documents rankings by various demographic variables


```{r}
purpose_of_ethics_doc_wide_demo <- 
purpose_of_ethics_doc %>% 
  mutate(id = 1:n()) %>% 
  left_join(purpose_of_ethics_doc_wide) %>% 
    mutate(rank = parse_number(variable)) 

names(purpose_of_ethics_doc_wide_demo)[6:12] <- 
  c("Work setting",
    "Ethnicity",
    "LGBTQIA+",
    "Gender",
    "Age",
    "Country of residence",
    "Geographic origin"
  )

# for each demo var
purpose_of_ethics_doc_wide_demo_wide <- 
purpose_of_ethics_doc_wide_demo %>% 
  pivot_longer(cols = `Work setting`:`Geographic origin`,
    names_to = "var1",
    values_to = "val1")

purpose_of_ethics_doc_wide_demo_wide_means <- 
purpose_of_ethics_doc_wide_demo_wide %>% 
  group_by(value, var1, val1) %>% 
  summarise(mean_rank = mean(rank)) 

purpose_of_ethics_doc_wide_demo_wide_means_plots <- 
purpose_of_ethics_doc_wide_demo_wide_means %>% 
  ungroup %>% 
  nest_by(var1) %>% 
  mutate(plot_out = list(ggplot(data = data) +
  aes(value, 
      mean_rank,
      group = val1,
      colour = val1) +
  geom_point(size = 4) +
  geom_line() +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_color_viridis_d(name = var1) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_y_reverse() ))

library(cowplot)
plot_grid(plotlist = purpose_of_ethics_doc_wide_demo_wide_means_plots$plot_out,
          align = "hv")

ggsave(here::here(paste0('figures/purpose-of-doc-demographics.png')),
       height = 10,
       width = 17)

```



Q: What type of ethical document(s) do you feel would best serve the SAA membership? 

Q: Are there additional ethical issues that should be addressed by the Principles? 
